\plaintitle

### 2016-09-19 drafts/labor/complexity.tex
pushing whatever i have...

% \msb{this paragraph needs a topic sentence. it's a wandering paragraph currently. is the claim here that evaluation is the limiting factor?}
\topic{Determining appropriate pay rates, informed by the careful measurement of workers,
isn't the only major challenge piecework faced; evaluation proved a limiting factor as well.}
\citeauthor{10.2307/23702539} enumerates some of the roles required
to facilitate piecework in the early \nth{20} century --- among them
    ``\dots~piecework clerks, inspectors, and `experts'\dots''
\cite{10.2307/23702539}.
% \citeauthor{10.2307/23702539} and \citeauthor{hart2016rise} may seem
% to be making differing claims about the limitations of piecework, but
% we argue that \citeauthor{10.2307/23702539}
% is simply making a more concrete observation illustrating the insight that
% \citeauthor{hart2016rise} later makes.
\citeauthor{10.2307/23702539} recognizes that
it's necessary for a successful piecework shop to employ
clerks,
inspectors, and
other experts to properly design and evaluate complex work.
\citeauthor{hart2016rise} later makes a more concrete observation of this hurdle,
as he argues an ultimate limit to how far this can go;
at some point, evaluating multidimensional work output for quality
(rather than for quantity) becomes infeasible.
In his words,
``if the quality of the output is more difficult to measure than the quantity,
perhaps because of `difficult--to--observe' production techniques,
then a piecework system is likely to encourage
an over--emphasis on quantity produced and an under-emphasis on quality''
\cite{hart2016rise}.
This, \citeauthor{hart2016rise} argues, may have fundamentally hamstrung piecework,
and ultimately precipitated its downfall,
especially with the increasing complexity of manufacturing work
over the course of the \nth{20} century.
\ali{just to be clear, this whole point is queuing up so that later I can be like
``Oh hey computers are terrific at evaluating some stuff really quickly.
And researchers in crowd work have done some work on workers evaluating other workers
(e.g. Find--Fix--Verify, arguably \textit{PeerStudio} does something along this line) \cite{bernsteinSoylent,Kulkarni:2015:PRP:2724660.2724670}''}

\topic{This isn't to say that complex work is outside of the realm of piecework;
indeed, we've discussed complex applications of crowd work already.}
As \citeauthor{hart2013rise} described,
the 1930s saw a flourishing of ingenious piecework design
almost strictly out of necessity ---
it was infeasible to provide new workers with
the comprehensive education that was familiar to men through apprenticeships
\cite{hart2013rise}.
This constraint led to much more tightly scoped work, and
(perhaps surprisingly at the time) more efficient allocations of workers,
who could now specialize in extremely narrowly defined roles.
The same could be said of Airy and his \textit{computers}
--- young boys whose preparations consisted principally of
a relatively specific mathematics curriculum
\cite{grier2013computers}.
Instead, the literature suggests that
piecework is tightly limited only when the application of piecework follows
a direct, perhaps even unimaginative,
mapping from an time--based regime to an output--based one.
When the work is redesigned from the ground up 
--- as we see with mathematicians in the \nth{19} century and
with the metalworking industry during the Second World War ---
it seems that we don't yet know the limits of complexity with regard to piecework.
% \msb{so what are we to take out of this? that there is no limit? just our creativity?}
\ali{This is \textit{almost} making a prediction,
but I wanted to take something punchy away from the stuff I cite here.}

\topic{Piecework researchers also argue that,
in addition to constraints on the kind of \textit{work} that's amenable to piecework,
only certain kinds of \textit{organizations} are amenable to piecework.}
% \msb{I think a reader wouldn't understand why that claim is relevant.
% Let's invert it, something like how
% the researchers point out that
% only certain kinds of organizations can effectively make use of piecework.}
\citeauthor{Brown01041990} discusses the organizational factors necessary for piecework to thrive,
arguing that piecework
``\dots~ is less likely in jobs with
a variety of duties than in jobs with a narrow set of routinized duties''
\cite{Brown01041990}
% \msb{that's confusing, is incentive pay the same as piecework?
% can you cut the first part of the quote and
% give us the noun instead of forcing us to figure it out?}.
\citeauthor{10.2307/23702539} adds further, that
successful cases of piecework owed themselves in part to the fact that
    ``\dots~only [the largest and most wealthy railroads] had the resources to~\dots
    pay the overhead involved in installing work reorganization''
\cite{10.2307/23702539}.
Together, \citeauthor{10.2307/23702539} and \citeauthor{Brown01041990}
make a persuasive argument that piecework is limited in complexity by
the capacity to endure managerial overhead while transitioning to a new system.
\msb{wait, wasn't that in a previous paragraph?
that should be joined with the managerial overhead text above}
\ali{I'm trying to make this argument that \cite{10.2307/23702539} is making two claims;
the first is that there are limits on doing hard, ``difficult--to--observe'' work,
and the second is that the cost propagates up to management,
putting limits on who can run piecework systems.
So\dots\\
\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Rewrite\\
\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Refactor\\
\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Drop}

\topic{There are other characteristics to effective complex piecework institutions, such as
appropriately designed management practices.}
\citeauthor{10.2307/2118435} describe the role of the foreman in West Virginia coal mines under the piecework model:
``The foreman had the power to hire and fire workers and allocate workplaces,
but then left the face--worker largely free to his own efforts so that
often he went all day without seeing the foreman''
\cite{10.2307/2118435}.
The general approach adopted by these West Virginia mines was,
as in other factories with active foremen,
to let the foreman be the intermediary between management and the worker.
Specifically, foremen were responsible for allocating resources and
understanding when and how to modify work as necessary
\cite{wray1949marginal}. \msb{Currently unclear what I'm supposed to take out of this. Summarize for me. What does this say about the complexity limits of piecework?}

\subsubsubsection{\whatchanged}
\msb{Before you get into this, summarize what I am supposed to have learned from the prior section on piecework. You're about to draw on those points to make your argument, so they need to be at the top of my mind here.}
\topic{Digital media have expanded the scope of
viable piecework by pushing drastically on the limits
cited by piecework researchers.}
The research on piecework tells us that
we should expect piecework to thrive in industries where
the nature of the work is limited in complexity
\cite{Brown01041990} \msb{I don't recall this point in the prior section: I remember management overhead and fixed costs of materials, I don't remember a paragraph about complexity. If we want to draw on that point, make it in the earlier section}.
Given the flourishing of on--demand labor platforms such as
Uber, AMT, and others, we ask ourselves
what --- if anything --- has changed.
We argue that
the Internet has trivialized
the costs and challenges of the earlier limiting factors for two reasons:
\begin{inlinelist}
  \item Technology make it much easier to do complex work aided by computers; and % is there a source for this? Is this necessary to cite?
  \item The Internet allows us to leverage the benefits of
        ``economies of scale'' at very little cost
        to the system--designer \cite{lessig2006code,miller2011understanding} \msb{I don't understand that argument. What is scaling here? The people? The tools for complex work?}
\end{inlinelist}.

>>>>

% \msb{this paragraph needs a topic sentence. it's a wandering paragraph currently. is the claim here that evaluation is the limiting factor?}
\topic{Determining appropriate pay rates, informed by the careful measurement of workers,
isn't the only major challenge piecework faced; evaluation proved a limiting factor as well.}
\citeauthor{10.2307/23702539} enumerates some of the roles required
to facilitate piecework in the early \nth{20} century --- among them
    ``\dots~piecework clerks, inspectors, and `experts'\dots''
\cite{10.2307/23702539}.
% \citeauthor{10.2307/23702539} and \citeauthor{hart2016rise} may seem
% to be making differing claims about the limitations of piecework, but
% we argue that \citeauthor{10.2307/23702539}
% is simply making a more concrete observation illustrating the insight that
% \citeauthor{hart2016rise} later makes.
\citeauthor{10.2307/23702539} further recognizes that
it's necessary for a successful piecework shop to employ
clerks,
inspectors, and
other experts to properly design and evaluate complex work.
\citeauthor{hart2016rise} later makes a more concrete observation of this hurdle,
as he argues an ultimate limit to how far this can go;
at some point, evaluating multidimensional work output for quality
(rather than for quantity) becomes infeasible.
In his words,
``if the quality of the output is more difficult to measure than the quantity,
perhaps because of `difficult--to--observe' production techniques,
then a piecework system is likely to encourage
an over--emphasis on quantity produced and an under-emphasis on quality''
\cite{hart2016rise}.
This, \citeauthor{hart2016rise} argues, may have fundamentally hamstrung piecework,
and ultimately precipitated its downfall,
especially with the increasing complexity of manufacturing work
over the course of the \nth{20} century.
\ali{just to be clear, this whole point is queuing up so that later I can be like
``Oh hey computers are terrific at evaluating some stuff really quickly.
And researchers in crowd work have done some work on workers evaluating other workers
(e.g. Find--Fix--Verify, arguably \textit{PeerStudio} does something along this line)
\cite{bernsteinSoylent,Kulkarni:2015:PRP:2724660.2724670}''}

\topic{The research seems to suggest that it was difficult to apply piecework to more skilled work,
particularly because maximizing on the advantages of piecework seemed to reward
smaller, more constrained, more narrowly trained tasks.}
For most of the \nth{19} century,
piecework was applied almost exclusively to farm and textile work.
% \citeauthor{grier2013computers} describes Airy's training and employment of young men to compute astronomical observations,
% mentioning that boys needed only to have ``the basic skills of mathematics''
% \cite{grier2013computers}.
% \citeauthor{hart2013rise} argue that piecework flourished in the first half of the \nth{20} century because
% tasks were decomposed --- a discussion we will return to later ---
% such that untrained workers could be trained relatively quickly on a very narrowly designed task
% \cite{hart2013rise}.
Work was simple and widely understood
--- farm workers didn't need to be trained on how to plow fields, or birth foals;
seamstresses knew how to sew together denim
\cite{10.2307/2338394,riisOtherSideLives}.
% The general argument for decomposition, we will see later,
% hinges on the promise that
% tasks can be \textit{made} less complex and time--consuming,
% if they're not already so.
% \ali{this section is here because i later talk about the complexity of work, and
% I want to keep the whole thing about how Uber, Google Maps, etc\dots make it possible to
% \begin{inlinelist}
% \item train workers more quickly (honestly just transmitting instructions over the Internet is so much quicker than what Airy had), and
% \item give workers necessary tools/information to get work done
% (e.g. ``Waze knows you should avoid the intersection 3 blocks down the road because of an accident''
% and (maybe) ``not great at math? WolframAlpha can confirm or just give you your answer'')
% \end{inlinelist}.}

\topic{This isn't to say that complex work is outside of the realm of piecework;
indeed, we've discussed complex applications of crowd work already.}
% As \citeauthor{hart2013rise} described,
While \citeauthor{hart2013rise} described a flourishing of ingenious piecework design,
much of it arose out of necessity ---
% almost strictly out of necessity ---
it was infeasible to provide new workers with
the comprehensive education that was familiar to men through apprenticeships
\cite{hart2013rise}.
While this constraint led to much more tightly scoped work
% (perhaps surprisingly at the time) more efficient allocations of workers,
who now had to specialize in extremely narrowly defined roles.
The same could be said of Airy and his \textit{computers}
--- young boys whose preparations consisted principally of
a relatively specific mathematics curriculum
\cite{grier2013computers}.
Instead, we argue that the literature suggests that
piecework is tightly limited only when the application of piecework follows
a direct, perhaps even unimaginative,
mapping from an time--based regime to an output--based one.
When the work is redesigned from the ground up 
--- as we see with mathematicians in the \nth{19} century and
with the metalworking industry during the Second World War ---
it seems that we don't yet know the limits of complexity with regard to piecework.
% \msb{so what are we to take out of this? that there is no limit? just our creativity?}
\ali{This is \textit{almost} making a prediction,
but I wanted to take something punchy away from the stuff I cite here.}

\topic{Piecework researchers also argue that,
in addition to constraints on the kind of \textit{work} that's amenable to piecework,
only certain kinds of \textit{organizations} are amenable to piecework.}
% \msb{I think a reader wouldn't understand why that claim is relevant.
% Let's invert it, something like how
% the researchers point out that
% only certain kinds of organizations can effectively make use of piecework.}
\citeauthor{Brown01041990} discusses the organizational factors necessary for piecework to thrive,
arguing that piecework
``\dots~ is less likely in jobs with
a variety of duties than in jobs with a narrow set of routinized duties''
\cite{Brown01041990}
% \msb{that's confusing, is incentive pay the same as piecework?
% can you cut the first part of the quote and
% give us the noun instead of forcing us to figure it out?}.
\citeauthor{10.2307/23702539} adds further, that
successful cases of piecework owed themselves in part to the fact that
    ``\dots~only [the largest and most wealthy railroads] had the resources to~\dots
    pay the overhead involved in installing work reorganization''
\cite{10.2307/23702539}.
Together, \citeauthor{10.2307/23702539} and \citeauthor{Brown01041990}
make a persuasive argument that piecework is limited in complexity by
the capacity to endure managerial overhead while transitioning to a new system.
\msb{wait, wasn't that in a previous paragraph?
that should be joined with the managerial overhead text above}
\ali{I'm trying to make this argument that \cite{10.2307/23702539} is making two claims;
the first is that there are limits on doing hard, ``difficult--to--observe'' work,
and the second is that the cost propagates up to management,
putting limits on who can run piecework systems.
So\dots\\
\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Rewrite\\
\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Refactor\\
\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Drop

Maybe I should just merge it with the next paragraph?}

\topic{There are other characteristics to effective complex piecework institutions, such as
appropriately designed management practices.}
\citeauthor{10.2307/2118435} describe the role of the foreman in West Virginia coal mines under the piecework model:
``The foreman had the power to hire and fire workers and allocate workplaces,
but then left the face--worker largely free to his own efforts so that
often he went all day without seeing the foreman''
\cite{10.2307/2118435}.
The general approach adopted by these West Virginia mines was,
as in other factories with active foremen,
to let the foreman be the intermediary between management and the worker.
Specifically, foremen were responsible for allocating resources and
understanding when and how to modify work as necessary
\cite{wray1949marginal}.
% \msb{Currently unclear what I'm supposed to take out of this. Summarize for me. What does this say about the complexity limits of piecework?}
The management of pieceworkers %, \citeauthor{10.2307/2118435} and later \citeauthor{wray1949marginal} argue,
demanded people in positions akin to foremen
--- intermediate managers people who were
\begin{inlinelist}
  \item familiar with and even sympathetic to the needs of workers,
  \item empowered by higher level management to make decisions, and
  \item relaxed enough in day--to--day work to allow workers to go about their work
\end{inlinelist}
\cite{wray1949marginal,10.2307/2118435}.

\subsubsubsection{\whatchanged}
\msb{Before you get into this, summarize what I am supposed to have learned from the prior section on piecework.
You're about to draw on those points to make your argument, so they need to be at the top of my mind here.}
\topic{Piecework makes a number of observations leading to the conclusion that
piecework's complexity is fundamentally bounded by several limitations,
chief among them the costs of managerial overhead and the transition thereto.}
\citeauthor{Brown01041990} and \citeauthor{10.2307/23702539}'s claims that
organizations can't adopt piecework unless
they're sufficiently large to absorb the cost of transitioning to a piecework system;
\citeauthor{10.2307/2118435} and \citeauthor{wray1949marginal}'s observations for
the importance of competent, effective managerial oversight
--- a human resource, which made the scaling cost prohibitively expensive for many
% in their times (\citeyear{10.2307/2118435} and \citeyear{wray1949marginal}, respectively)
\cite{10.2307/2118435,wray1949marginal,10.2307/23702539,Brown01041990}.
\ali{Something like this? Should I dig deeper?
I can line these points up in a way to make the next paragraph sort of obvious or inevitably,
depending on how actively engaged someone is while reading this\dots}

\topic{Digital media have expanded the scope of
viable piecework by pushing drastically on the limits
cited by piecework researchers.}
The research on piecework tells us that
we should expect piecework to thrive in industries where
the nature of the work is limited in complexity
\cite{Brown01041990}.
% \msb{I don't recall this point in the prior section:
% I remember management overhead and fixed costs of materials,
% I don't remember a paragraph about complexity.
% If we want to draw on that point, make it in the earlier section}
Given the flourishing of on--demand labor platforms such as
Uber, AMT, and others, we ask ourselves
what --- if anything --- has changed.
We argue that
the Internet has trivialized
the costs and challenges of the earlier limiting factors %  for two reasons:
% \begin{inlinelist}
  % \item 
  because technology make it easier
  \begin{inlinelist}
  \item to do complex work aided by computers and
  \item to evaluate and manage workers as they do increasingly complex work,
        even observing their work to an otherwise unprecedented granularity.
  \end{inlinelist}

%   \item The Internet allows us to leverage the benefits of
%         ``economies of scale'' at very little cost
%         to the system--designer \cite{lessig2006code,miller2011understanding}.% \msb{I don't understand that argument. What is scaling here? The people? The tools for complex work?}
% \end{inlinelist}

### 2016-09-19 drafts/labor/introduction.tex
pushing whatever i have...

\topic{The past decade has seen a flourishing of \textit{on--demand work},
largely driven by the reformulation of work as
the constituent parts of larger tasks.}
This framing of work into de-contextualized, modular blocks
enables computation to hire workers at scale through open calls on the internet~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed paid participants then engage in the work whenever their schedules allow,
with little to no awareness of the broader context of the work, and 
with (often) fleeting identities and associations
\cite{martin2014being,uberAlgorithm}.
In this paper, we use the term on--demand work to join a pair of related phenomena:%We've seen on--demand work passing by a number of aliases:
(1) \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work; and
(2) \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.
%as the \textit{sharing economy} --- where participants can rent out their valuables (such as a car, a home, etc\dots)
%on an ad hoc basis when they don't need it.
The realization that complex tasks can be accomplished by directing and managing these crowds of workers spurred industry to flock to sites of labor
like AMT and Uber to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems and designs that enable new forms of production~(e.g., \cite{bernsteinSoylent,vizwiz,paolacci2010running}).

@@

\topic{This body of research has sought to answer one central question:
What does the future hold for on--demand work and those that do it?}
Researchers have offered their input on this open question along three major threads:
\begin{enumerate}
  \item \nameref{sec:complexity}?
        Specifically,
        \begin{inlinelist}
          \item How complex are the goals that crowd work can accomplish?, and
          \item What kinds of goals and industries may eventually utilize it?
        \end{inlinelist}
        \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
              Nebeling:2016:WCW:2858036.2858169,
              Hahn:2016:KAB:2858036.2858364};
  \item \nameref{sec:decomposition}?
        \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
              LykourentzouPersonalityMatters,KucherbaevReLauncher,
              Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
              Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
  \item \nameref{sec:relationships}?
        \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
              whyWouldAnyoneBrewer,takingAHITMcInnis}
        % \ali{is this about collective action? yes; also, career growth? education? governance? sociality? wages?}
\end{enumerate}
This research literature has largely sought to answer these questions by examining the present phenomenon.
So far, it has not offered a framing for holistically explaining
the developments in worker processes that researchers have developed, or
the emergent phenomena in social environments;
nor has any research, to our knowledge,
gone as far as predict future developments.

>>>>

\topic{The past decade has seen a flourishing of \textit{on--demand work},
largely driven by the reformulation of work as
the constituent parts of larger tasks.}
This framing of work into de--contextualized, modular blocks
enables computation to hire workers at scale through open calls on the Internet
\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed paid participants then engage in the work whenever their schedules allow,
with little to no awareness of the broader context of the work, and 
with (often) fleeting identities and associations
\cite{martin2014being,uberAlgorithm}.
In this paper, we use the term on--demand work to join a pair of related phenomena:
\begin{inlinelist}
\item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work; and
\item \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.
\end{inlinelist}
The realization that complex tasks can be accomplished by directing and managing these crowds of workers spurred industry to flock to sites of labor
like AMT and Uber to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems and designs that enable new forms of production
(e.g., \cite{bernsteinSoylent,vizwiz,paolacci2010running}).

@@

\topic{This body of research has sought to answer one central question:
What does the future hold for on--demand work and those that do it?}
Researchers have offered their input on this open question along three major threads:
\begin{enumerate}
  \item \nameref{sec:complexity}?
        Specifically,
        \begin{inlinelist}
          \item How complex are the goals that crowd work can accomplish?, and
          \item What kinds of goals and industries may eventually utilize it?
        \end{inlinelist}
        \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
              Nebeling:2016:WCW:2858036.2858169,
              Hahn:2016:KAB:2858036.2858364};
  \item \nameref{sec:decomposition}?
        \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
              LykourentzouPersonalityMatters,KucherbaevReLauncher,
              Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
              Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
  \item \nameref{sec:relationships}?
        \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
              whyWouldAnyoneBrewer,takingAHITMcInnis}
\end{enumerate}
This research literature has largely sought to answer these questions by examining the present phenomenon.
So far, it has not offered a framing for holistically explaining
the developments in worker processes that researchers have developed, or
the emergent phenomena in social environments;
nor has any research, to our knowledge,
gone as far as predict future developments.

### 2016-09-19 drafts/labor/piecework_lit.tex
pushing whatever i have...

\documentclass[trackingWork]{subfiles}

% \renewcommand{\topic}[1]{#1}

\makeatletter
\makeatother
\begin{document}
\section{A Review of Piecework}
\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowd work
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the subsequent sections --- and indeed, the entire argument.
Specifically, we will
\begin{inlinelist}
  \item define ``piecework'' as researchers in the topic understood it;
  \item trace the rise of piecework at a very high level,
        identifying key figures and ideas during this time; and finally
  \item look at the fall of piecework, such as it was,
        considering in particular
        the factors that may have led to piecework's
        \msb{(and crowd work's?)}
        \ali{I wanted to leave the door open later to argue that
        crowd work can change its trajectory and \textit{not have a demise}.
        Is that too subtle here? I could have an aside like
        ``(and help crowd work avoid the same fate)''. Thoughts?}
        eventual demise.
\end{inlinelist}

\subsection{What was piecework?}
\topic{While ``piecework'' has proven difficult to concretize from the literature,
we can trace a constellation of characteristics of piece work that recur throughout the literature.}
We'll follow the history of research, collecting
descriptions,
examples, and
provided definitions of piecework, trying
to trace the outline of a working understanding of
\textit{what piecework is}.

@@

\topic{\citeauthor{hart2013rise} offer another series of compelling insights toward the question of the features that sprout from piecework.}
In their reflection on the features endemic to piecework in the 1930s,
which they describe as the ``heyday'' of piecework's prominence;
among them were the following:
\begin{inlinelist}
\item ``female workers who generally had less training'' had to be trained in narrower subsets of the general body of skills that conventional (male) apprentices would undertake, and
\item workers with specific slices of skills could be more appropriately matched to suitable tasks
\end{inlinelist}
\cite{hart2013rise}.
Piecework thus opened the door for people who previously couldn't participate in the labor market
--- either for lack of training or for other reasons ---
to do so, and to acquire job skills incrementally.
Workers without conventional training
--- like women, who had no such opportunities to engage in engineering and metalworking apprenticeships as men did ---
could be trained very narrowly on a very tightly constrained task,
demonstrate proficiency, and become experts in their own ways.
% \msb{explain that in your own words now --- that the piecework infrastructure allows lower-skilled people to be used for skilled positions, by building the expertise into the workflow}

% \topic{Consolidating what we've learned from these sources about piecework, we might be able to arrive at a working definition that will
% suffice for our needs.} \msb{Again, I don't think the goal for the reader should be to arrive at a definition through a narrative, they instead need to know how to think about piecework. This  paragraph could probably be cut}
% Those needs being that the definition be
% \begin{inlinelist}
% \item faithful to the historical cases of piecework that we see in the scholarship; and
% \item relevant and informative to potential cases of piecework today.
% \end{inlinelist}
% We offer the following:
% that \pieceworkdefinition.

@@

\ali{I wanted that paragraph to be about the rising popularity and application of piecework
(especially as it approached its ``heyday'' (I like \cite{hart2013rise})),
coming from humble beginnings as it found its footing.
Given that intent\dots~

@@

% \topic{Piecework took a circuitous path in its rise to the mainstream%
% each time finding additional ways to leverage the advantages of piecework.}\msb{that sentence has some circular logic going on, it took a path to leveraging the advantages of itself?} \ali{Yeah, this is really weird; commenting out for now.}
First applied to farm work, as
\citeauthor{hughRaynbirdTaskWork} and others illustrate,
the practice remained relatively obscure until
it was brought to the textile industry
\cite{hughRaynbirdTaskWork}.
At the turn of the \nth{20} century,
when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
\citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
offering guidance on how best to manage pieceworkers
\cite{riisOtherSideLives,norton1900textile}.
\citeauthor{clark1908cotton}, for instance,
relays his observations of textile mill pieceworkers and his realization that
``When he works by the day the Italian operative wishes to leave before the whistle blows,
but if he works by the piece he will work as many hours as it is possible for him to stand''
\cite{clark1908cotton}.
Best practices regarding the measurement and management of
piecework rates, and of workers in the engineering industry,
were beginning to take shape
\cite{burton1899commercial}.

@@

% \topic{Necessity drove the widespread adoption of piecework and work decomposition%
% , perhaps more even than Ford's assembly line.}

@@

\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed just a few decades earlier.}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
We summarize some of the salient suggestions here:
\begin{inlinelist}
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance;
\item the shifting of these industries (manufacturing, clothing, etc\dots)
to other countries;
\item the quality of ``multidimensional'' work becoming too difficult to evaluate.
\end{inlinelist}
\cite{hart2013rise}.
% the gendered and stratified dynamics at play in the matchgirls strike

% \topic{Piecework declined in popularity in the United States, but
% its spirit thrived through the mid--\nth{20} century in
% scientific management and Fordism.}
% Concurrent with the fall of piecework was arguably the rise of factories as centers of work.

\subsection{Why is piecework relevant to crowdwork?}
\msb{I assume this is forthcoming?}
Using the definition of piecework that we came up with earlier, we argue that
crowd work is fundamentally an instantiation of piecework, and
that we can more precisely anticipate the answers to the open research questions we discussed earlier.
We'll show that the dimensions of crowd work that the broader HCI community has been studying
align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.

>>>>

\documentclass[trackingWork]{subfiles}
\makeatletter
\makeatother
\begin{document}

\section{A Review of Piecework}

\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowd work
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the subsequent sections --- and indeed, the entire argument.
Specifically, we will
\begin{inlinelist}
  \item define ``piecework'' as researchers in the topic understood it;
  \item trace the rise of piecework at a very high level,
        identifying key figures and ideas during this time; and finally
  \item look at the fall of piecework, such as it was,
        considering in particular
        the factors that may have led to piecework's
        \msb{(and crowd work's?)}
        \ali{I wanted to leave the door open later to argue that
        crowd work can change its trajectory and \textit{not have a demise}.
        Is that too subtle here? I could have an aside like
        ``(and help crowd work avoid the same fate)''. Thoughts?}
        eventual demise.
\end{inlinelist}

\subsection{What was piecework?}

\topic{While ``piecework'' has proven difficult to concretize from the literature,
we can trace a constellation of characteristics of piece work that recur throughout the literature.}
We'll follow the history of research, collecting
descriptions,
examples, and
provided definitions of piecework, trying
to trace the outline of a working understanding of
\textit{what piecework is}.

@@

\topic{\citeauthor{hart2013rise} offer another series of compelling insights toward the question of the features that sprout from piecework.}
In their reflection on the features endemic to piecework in the 1930s,
which they describe as the ``heyday'' of piecework's prominence;
among them were the following:
\begin{inlinelist}
\item ``female workers who generally had less training'' had to be trained in narrower subsets of the general body of skills that conventional (male) apprentices would undertake, and
\item workers with specific slices of skills could be more appropriately matched to suitable tasks
\end{inlinelist}
\cite{hart2013rise}.
Piecework thus opened the door for people who previously couldn't participate in the labor market
--- either for lack of training or for other reasons ---
to do so, and to acquire job skills incrementally.
Workers without conventional training
--- like women, who had no such opportunities to engage in engineering and metalworking apprenticeships as men did ---
could be trained very narrowly on a very tightly constrained task,
demonstrate proficiency, and become experts in their own ways.

@@

\ali{I wanted that paragraph to be about the rising popularity and application of piecework
(especially as it approached its ``heyday'' (a la \cite{hart2013rise})),
coming from humble beginnings as it found its footing.
Given that intent\dots~

@@

\topic{First applied to farm work, as
\citeauthor{hughRaynbirdTaskWork} and others illustrate,
the practice remained relatively obscure until
it was brought to the textile industry
\cite{hughRaynbirdTaskWork}.}
At the turn of the \nth{20} century,
when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
\citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
offering guidance on how best to manage pieceworkers
\cite{riisOtherSideLives,norton1900textile}.
\citeauthor{clark1908cotton}, for instance,
relays his observations of textile mill pieceworkers and his realization that
``When he works by the day the Italian operative wishes to leave before the whistle blows,
but if he works by the piece he will work as many hours as it is possible for him to stand''
\cite{clark1908cotton}.
Best practices regarding the measurement and management of
piecework rates, and of workers in the engineering industry,
were beginning to take shape
\cite{burton1899commercial}.

@@

@@

\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed just a few decades earlier.}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
We summarize some of the salient suggestions here:
\begin{inlinelist}
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance;
\item the shifting of these industries (manufacturing, clothing, etc\dots)
to other countries;
\item the quality of ``multidimensional'' work becoming too difficult to evaluate.
\end{inlinelist}
\cite{hart2013rise}.

\subsection{Why is piecework relevant to crowdwork?}

\msb{I assume this is forthcoming?}
Using the definition of piecework that we came up with earlier, we argue that
crowd work is fundamentally an instantiation of piecework, and
that we can more precisely anticipate the answers to the open research questions we discussed earlier.
We'll show that the dimensions of crowd work that the broader HCI community has been studying
align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.
